---
title: "The anatomy of a paired count"
description: >
  What a pair of counts is made of, and the two coordinate systems you can
  describe it in — interactively, with every knob linked to every other.
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Every joint family in this package — `bipois()`, `binegbin()`,
`binegbin_joint()` — is built by **trivariate reduction**. Three independent
counts are drawn, and the two observed counts share one of them:

$$
y_{\text{em}} = N_{\text{shared}} + N_{10}
\qquad
y_{\text{lb}} = N_{\text{shared}} + N_{01}
$$

That shared term is the whole point. It is what makes the pair correlated, it
is never observed, and it is marginalised out analytically in the likelihood.
Everything else is the part each source saw alone.

The widget below lets you take that apart. The novel bit is that it exposes
**two coordinate systems at once, each driving the other**.

## Why two coordinate systems

The dpars a family actually takes are three rates — `mu` for the shared
component, `lambdaem` and `lambdalb` for the two excesses. That is what the
likelihood consumes, but it is awkward to reason in: raise the overall level of
counting and all three move together, so no single one of them answers "how
much was there", "how much did the two sources agree", or "which source ran
high".

Those three questions have their own coordinates:

| | | |
|---|---|---|
| $M$ | overall level | $\mu + (\lambda_{\text{em}} + \lambda_{\text{lb}})/2$ |
| $f$ | congruence — the share of $M$ both sources saw | $\mu / M$ |
| $\beta$ | method bias, bounded on $[-1,1]$ | $(\lambda_{\text{em}} - \lambda_{\text{lb}})/(\lambda_{\text{em}} + \lambda_{\text{lb}})$ |

The map between them is a bijection, so neither set is more "real" — they are
two descriptions of one object. Drag anything below and watch the other five
respond.

```{=html}
<style>
  .pca { --pca-shared:#8A8072; --pca-em:#2D6A7F; --pca-lb:#C4622D;
         --pca-ink:#3A3A3A; --pca-shared-d:#6B5D4F; --pca-rule:#999;
         --pca-panel:rgba(127,127,127,.06); --pca-edge:rgba(127,127,127,.35);
         font-size:14px; }
  @media (prefers-color-scheme: dark) {
    .pca { --pca-shared:#B3A692; --pca-em:#5FA8C4; --pca-lb:#E08A50;
           --pca-ink:#D8D4CE; --pca-shared-d:#C0B49E; --pca-rule:#777; }
  }
  :root[data-theme="dark"] .pca {
    --pca-shared:#B3A692; --pca-em:#5FA8C4; --pca-lb:#E08A50;
    --pca-ink:#D8D4CE; --pca-shared-d:#C0B49E; --pca-rule:#777;
  }
  :root[data-theme="light"] .pca {
    --pca-shared:#8A8072; --pca-em:#2D6A7F; --pca-lb:#C4622D;
    --pca-ink:#3A3A3A; --pca-shared-d:#6B5D4F; --pca-rule:#999;
  }
  .pca-cols { display:grid; grid-template-columns:repeat(auto-fit,minmax(300px,1fr));
              gap:1.4rem; margin:.6rem 0 1rem; }
  .pca-group h4 { margin:.1rem 0 .5rem; font-size:.95rem; letter-spacing:.02em; }
  .pca-group { border:1px solid var(--pca-edge); border-radius:7px; padding:.7rem .85rem; }
  .pca-dial { display:flex; align-items:center; gap:.5rem; margin:.3rem 0; }
  .pca-dial .lab { flex:0 0 8.5em; font-size:.86rem; }
  .pca-dial .lab sub { line-height:0; font-size:.75em; }
  .pca-dial input[type=range] { flex:1; min-width:0; accent-color:#7A8288; }
  .pca-dial .val { flex:0 0 4.6em; font-variant-numeric:tabular-nums;
    text-align:right; font-size:.83rem; color:var(--pca-ink);
    background:transparent; border:1px solid var(--pca-edge);
    border-radius:4px; padding:.05rem .25rem; font-family:inherit; }
  .pca-note { font-size:.8rem; opacity:.72; margin:.45rem 0 0; line-height:1.45; }
  .pca-bar { display:flex; gap:.7rem; align-items:center; flex-wrap:wrap; margin:.2rem 0 .7rem; }
  .pca-plot { overflow-x:auto; }
  .pca svg text { fill:currentColor; }
  .pca-flag { display:inline-block; font-size:.78rem; padding:.08rem .4rem;
    border-radius:3px; background:rgba(200,120,40,.16); margin-left:.3rem; }
</style>

<div class="pca" id="pca">
  <div class="pca-bar">
    <button type="button" id="pca-resim">re-simulate</button>
    <label style="font-size:.82rem">seed
      <input type="number" id="pca-seed" value="7" style="width:5em"></label>
    <button type="button" id="pca-reset">reset</button>
  </div>

  <div class="pca-cols">
    <div class="pca-group">
      <h4>Interpretable coordinates</h4>
      <div id="pca-mfd"></div>
      <p class="pca-note" id="pca-note-mfd"></p>
    </div>
    <div class="pca-group">
      <h4>Native dpars (what <code>binegbin()</code> takes)</h4>
      <div id="pca-nat"></div>
      <p class="pca-note" id="pca-note-nat"></p>
    </div>
  </div>

  <div class="pca-plot" id="pca-plot"></div>
</div>

<script>
(function () {
  "use strict";

  // ---- seeded RNG, so a given seed always gives the same picture ----------
  var rngState = 1;
  function seedRng(s){ rngState = (s >>> 0) || 1; }
  function rand(){
    rngState |= 0; rngState = (rngState + 0x6D2B79F5) | 0;
    var t = Math.imul(rngState ^ (rngState >>> 15), 1 | rngState);
    t = (t + Math.imul(t ^ (t >>> 7), 61 | t)) ^ t;
    return ((t ^ (t >>> 14)) >>> 0) / 4294967296;
  }
  function randn(){
    var u = 0, v = 0;
    while (u === 0) u = rand();
    while (v === 0) v = rand();
    return Math.sqrt(-2 * Math.log(u)) * Math.cos(2 * Math.PI * v);
  }
  function rgamma(shape){
    if (shape < 1) return rgamma(shape + 1) * Math.pow(rand(), 1 / shape);
    var d = shape - 1 / 3, c = 1 / Math.sqrt(9 * d);
    for (;;) {
      var x = randn(), v = Math.pow(1 + c * x, 3);
      if (v <= 0) continue;
      var u = rand();
      if (u < 1 - 0.0331 * x * x * x * x) return d * v;
      if (Math.log(u) < 0.5 * x * x + d * (1 - v + Math.log(v))) return d * v;
    }
  }
  function rpois(l){
    if (l <= 0) return 0;
    if (l < 30) { var L = Math.exp(-l), k = 0, p = 1;
      do { k++; p *= rand(); } while (p > L); return k - 1; }
    return Math.max(0, Math.round(l + Math.sqrt(l) * randn()));
  }
  // NB2 with SD-scale kappa: phi = 1/kappa^2, and kappa = 0 IS Poisson.
  function rnbinom(mu, kappa){
    if (kappa <= 0) return rpois(mu);
    var phi = 1 / (kappa * kappa);
    return rpois(rgamma(phi) * (mu / phi));
  }

  // ---- the map, stated once ----------------------------------------------
  // Mirrors binegbin_mfd_to_dpars() / binegbin_dpars_to_mfd() exactly.
  function toNative(s){
    var mid = s.M * (1 - s.f);
    return {
      mu:     s.M * s.f,
      lem:    mid * (1 + s.b),
      llb:    mid * (1 - s.b),
      shapes: s.kA <= 0 ? Infinity : 1 / (s.kA * s.kA),
      shapex: s.kX <= 0 ? Infinity : 1 / (s.kX * s.kX)
    };
  }
  // Degenerate directions: f undefined at M = 0, bias undefined when there is
  // no excess to be biased. Hold the last identified value rather than
  // snapping to 0 -- 0 would assert an unbiased method the state cannot
  // support, and would make the knob jump under the user's finger.
  var lastF = 0.67, lastB = 0, degenerate = null;
  function fromNative(n){
    var sum = n.lem + n.llb, M = n.mu + sum / 2;
    degenerate = null;
    var f = M > 0 ? n.mu / M : (degenerate = "M = 0: congruence undefined", lastF);
    var b = sum > 0 ? (n.lem - n.llb) / sum
                    : (degenerate = "no excess: bias unidentified", lastB);
    if (M > 0) lastF = f;
    if (sum > 0) lastB = b;
    return {
      M: M, f: f, b: b,
      kA: isFinite(n.shapes) ? 1 / Math.sqrt(n.shapes) : 0,
      kX: isFinite(n.shapex) ? 1 / Math.sqrt(n.shapex) : 0
    };
  }

  // ---- dial specs --------------------------------------------------------
  // Shape dials are log10-scaled: shape spans orders of magnitude, and its
  // Poisson limit is at infinity. LOG_MAX is the top of the slider; kappa = 0
  // parks it there and the readout shows the true infinity.
  var LOG_MIN = -1, LOG_MAX = 3;

  var MFD = [
    {k:"M",  lab:"Level <b>M</b>",             min:0.5, max:30, step:0.5,  dp:1},
    {k:"f",  lab:"Congruence <b>f</b>",        min:0,   max:1,  step:0.01, dp:2},
    {k:"b",  lab:"Bias <b>&beta;</b>",         min:-1,  max:1,  step:0.02, dp:2},
    {k:"kA", lab:"<b>&kappa;<sub>A</sub></b> shared",  min:0, max:3, step:0.05, dp:2},
    {k:"kX", lab:"<b>&kappa;<sub>X</sub></b> excess",  min:0, max:3, step:0.05, dp:2}
  ];
  var NAT = [
    {k:"mu",     lab:"<code>mu</code>",       min:0, max:30, step:0.1, dp:2},
    {k:"lem",    lab:"<code>lambdaem</code>", min:0, max:30, step:0.1, dp:2},
    {k:"llb",    lab:"<code>lambdalb</code>", min:0, max:30, step:0.1, dp:2},
    {k:"shapes", lab:"<code>shapes</code>",   log:true, dp:2},
    {k:"shapex", lab:"<code>shapex</code>",   log:true, dp:2}
  ];

  var DEFAULTS = {M:12, f:0.67, b:0, kA:0.6, kX:1.0};
  var S = Object.assign({}, DEFAULTS);
  var NSETS = 10, sets = [];

  var $ = function(id){ return document.getElementById(id); };
  var dials = {};

  function buildDial(spec, host, side){
    var wrap = document.createElement("div");
    wrap.className = "pca-dial";
    wrap.innerHTML = '<span class="lab">' + spec.lab + "</span>";
    var inp = document.createElement("input");
    inp.type = "range";
    if (spec.log) { inp.min = LOG_MIN; inp.max = LOG_MAX; inp.step = 0.01; }
    else { inp.min = spec.min; inp.max = spec.max; inp.step = spec.step; }
    var num = document.createElement("input");
    num.type = "text"; num.className = "val";
    wrap.appendChild(inp); wrap.appendChild(num);
    host.appendChild(wrap);
    dials[spec.k] = {inp:inp, num:num, spec:spec, side:side};

    // One side is the source of truth per interaction: the side you touched is
    // read in full, converted, and written to the other. Neither side mutates
    // the other incrementally, so repeated round-tripping cannot drift.
    inp.addEventListener("input", function(){ drive(side); });
    num.addEventListener("change", function(){
      var v = parseFloat(num.value);
      if (isFinite(v)) { setDial(spec.k, v); drive(side); }
      else render();
    });
  }

  function setDial(k, value){
    var d = dials[k];
    if (d.spec.log) {
      var lv = (!isFinite(value) || value <= 0) ? LOG_MAX
             : Math.min(LOG_MAX, Math.max(LOG_MIN, Math.log10(value)));
      d.inp.value = lv;
    } else {
      d.inp.value = Math.min(d.spec.max, Math.max(d.spec.min, value));
    }
  }
  function getDial(k){
    var d = dials[k];
    return d.spec.log ? Math.pow(10, parseFloat(d.inp.value)) : parseFloat(d.inp.value);
  }

  function drive(side){
    if (side === "mfd") {
      S = {M:getDial("M"), f:getDial("f"), b:getDial("b"),
           kA:getDial("kA"), kX:getDial("kX")};
      degenerate = null;
    } else {
      // kappa = 0 is unreachable from a finite shape slider; treat the top of
      // the log range as the Poisson limit so the two sides can agree there.
      var sh = parseFloat(dials.shapes.inp.value) >= LOG_MAX ? Infinity : getDial("shapes");
      var sx = parseFloat(dials.shapex.inp.value) >= LOG_MAX ? Infinity : getDial("shapex");
      S = fromNative({mu:getDial("mu"), lem:getDial("lem"),
                      llb:getDial("llb"), shapes:sh, shapex:sx});
    }
    resimulate();
  }

  function resimulate(){
    seedRng(parseInt($("pca-seed").value, 10) || 1);
    var n = toNative(S), out = [];
    for (var i = 0; i < NSETS; i++) {
      var A = rnbinom(n.mu, S.kA);
      out.push({A:A,
                em:A + rnbinom(n.lem, S.kX),
                lb:A + rnbinom(n.llb, S.kX)});
    }
    sets = out;
    render();
  }

  function fmt(v, dp){
    if (!isFinite(v)) return "∞";
    return v.toFixed(dp);
  }

  function render(){
    var n = toNative(S);
    // Write BOTH sides from the single state, every time.
    setDial("M", S.M); setDial("f", S.f); setDial("b", S.b);
    setDial("kA", S.kA); setDial("kX", S.kX);
    setDial("mu", n.mu); setDial("lem", n.lem); setDial("llb", n.llb);
    setDial("shapes", n.shapes); setDial("shapex", n.shapex);

    dials.M.num.value  = fmt(S.M, 1);
    dials.f.num.value  = fmt(S.f, 2);
    dials.b.num.value  = fmt(S.b, 2);
    dials.kA.num.value = fmt(S.kA, 2);
    dials.kX.num.value = fmt(S.kX, 2);
    dials.mu.num.value     = fmt(n.mu, 2);
    dials.lem.num.value    = fmt(n.lem, 2);
    dials.llb.num.value    = fmt(n.llb, 2);
    dials.shapes.num.value = fmt(n.shapes, 2);
    dials.shapex.num.value = fmt(n.shapex, 2);

    $("pca-note-mfd").innerHTML =
      "&kappa; = 0 is the Poisson floor, reachable at a finite zero."
      + (degenerate ? '<span class="pca-flag">' + degenerate + " &mdash; holding last value</span>" : "");
    $("pca-note-nat").innerHTML =
      "<code>shapes</code>/<code>shapex</code> are NB2 <i>&phi;</i>: <b>larger means less "
      + "overdispersed</b>, and Poisson is the <i>&phi;</i> &rarr; &infin; limit &mdash; off the end "
      + "of any finite slider. That asymmetry is why &kappa; = 1/&radic;<i>&phi;</i> is the nicer dial.";

    $("pca-plot").innerHTML = draw(n);
  }

  // ---- drawing -----------------------------------------------------------
  var SLOT = 12, GW = 11, COL = 15, PAIRW = 2 * COL, PITCH = PAIRW + 15;
  var ML = 46, MR = 14, MT = 16, MB = 32, BARW = 17, BARGAP = 5, BOXPAD = 9;
  var PANELW = 2 * BARW + BARGAP + 2 * BOXPAD;

  function yMaxOf(n){
    var vA = n.mu + S.kA * S.kA * n.mu * n.mu;
    var need = Math.max(n.lem, n.llb) * 0 + Math.max(
      (n.mu + n.lem) + 3 * Math.sqrt(vA + n.lem + S.kX * S.kX * n.lem * n.lem),
      (n.mu + n.llb) + 3 * Math.sqrt(vA + n.llb + S.kX * S.kX * n.llb * n.llb)
    );
    return Math.max(8, Math.min(34, Math.ceil(need)));
  }

  function draw(n){
    var yMax = yMaxOf(n), panelX = ML, setsX = panelX + PANELW + 20;
    var W = setsX + NSETS * PITCH - (PITCH - PAIRW) + MR;
    var H = MT + yMax * SLOT + MB;
    var Y = function(v){ return MT + (yMax - v) * SLOT; };
    var s = '<svg width="' + W + '" height="' + H + '" viewBox="0 0 ' + W + " " + H
          + '" font-size="10" role="img" aria-label="Decomposition of a paired count into a shared component and two source-specific excesses">';

    // expected panel: CONTINUOUS bars, because these are rates, not counts
    var boxY = Y(yMax) - 4, boxH = Y(0) - boxY + 4;
    s += '<rect x="' + (panelX - 2) + '" y="' + boxY + '" width="' + (PANELW + 4)
       + '" height="' + boxH + '" fill="var(--pca-panel)" stroke="var(--pca-edge)" rx="4"/>';
    s += '<text x="' + (panelX + PANELW / 2) + '" y="' + (boxY - 4)
       + '" text-anchor="middle" opacity=".7">expected</text>';

    var lam = [n.llb, n.lem], cols = ["var(--pca-lb)", "var(--pca-em)"], nm = ["LB", "EM"];
    for (var k = 0; k < 2; k++) {
      var bx = panelX + BOXPAD + k * (BARW + BARGAP);
      s += '<rect x="' + bx + '" y="' + Y(n.mu).toFixed(1) + '" width="' + BARW
         + '" height="' + (n.mu * SLOT).toFixed(1) + '" fill="var(--pca-shared)" fill-opacity=".45"/>';
      s += '<rect x="' + bx + '" y="' + Y(n.mu + lam[k]).toFixed(1) + '" width="' + BARW
         + '" height="' + (lam[k] * SLOT).toFixed(1) + '" fill="' + cols[k] + '" fill-opacity=".45"/>';
      s += '<text x="' + (bx + BARW / 2) + '" y="' + (Y(0) + 12)
         + '" text-anchor="middle" font-weight="600" fill="' + cols[k] + '">' + nm[k] + "</text>";
    }

    // the two parameter rules cross every simulated pair
    s += '<line x1="' + (panelX + BOXPAD) + '" y1="' + Y(S.M).toFixed(1) + '" x2="' + (W - MR + 6)
       + '" y2="' + Y(S.M).toFixed(1) + '" stroke="var(--pca-ink)" stroke-width="1.2" stroke-dasharray="1.5,2.5"/>';
    s += '<text x="' + (ML - 6) + '" y="' + (Y(S.M) + 3.5).toFixed(1)
       + '" text-anchor="end" font-style="italic" font-size="12">M</text>';
    s += '<line x1="' + (panelX + BOXPAD) + '" y1="' + Y(n.mu).toFixed(1) + '" x2="' + (W - MR + 6)
       + '" y2="' + Y(n.mu).toFixed(1) + '" stroke="var(--pca-shared-d)" stroke-width="1.2" stroke-dasharray="6,3"/>';
    s += '<text x="' + (ML - 6) + '" y="' + (Y(n.mu) + 3.5).toFixed(1)
       + '" text-anchor="end" font-size="12">mu</text>';

    // realised pairs: DISCRETE marks
    for (var i = 0; i < NSETS; i++) {
      var d = sets[i], pcx = setsX + i * PITCH + PAIRW / 2;
      var sides = [["lb", d.lb, "var(--pca-lb)"], ["em", d.em, "var(--pca-em)"]];
      for (var j = 0; j < 2; j++) {
        var side = sides[j], x = (j === 0) ? pcx - COL : pcx;
        var shown = Math.min(side[1], yMax);
        for (var q = 1; q <= shown; q++) {
          var col = (q <= d.A) ? "var(--pca-shared)" : side[2];
          s += '<rect x="' + (x + (COL - GW) / 2).toFixed(1) + '" y="' + (Y(q - 0.5) - GW / 2 + 1).toFixed(1)
             + '" width="' + GW + '" height="' + (GW - 2) + '" rx="2.5" fill="' + col + '"/>';
        }
        if (side[1] > yMax) {
          s += '<text x="' + (x + COL / 2).toFixed(1) + '" y="' + (Y(yMax) - 3).toFixed(1)
             + '" text-anchor="middle" opacity=".8">+' + (side[1] - yMax) + "</text>";
        }
      }
      s += '<line x1="' + (pcx - COL - 2).toFixed(1) + '" y1="' + Y(d.A).toFixed(1) + '" x2="'
         + (pcx + COL + 2).toFixed(1) + '" y2="' + Y(d.A).toFixed(1)
         + '" stroke="var(--pca-shared-d)" stroke-width="1.6"/>';
    }

    s += '<line x1="' + (panelX - 2) + '" y1="' + Y(0).toFixed(1) + '" x2="' + (W - MR + 6)
       + '" y2="' + Y(0).toFixed(1) + '" stroke="var(--pca-rule)"/>';
    s += '<text x="' + (ML - 6) + '" y="' + (Y(0) + 3.5).toFixed(1) + '" text-anchor="end" opacity=".6">0</text>';
    s += '<text x="' + (setsX + (NSETS * PITCH - (PITCH - PAIRW)) / 2) + '" y="' + (H - 4)
       + '" text-anchor="middle" opacity=".6">ten simulated pairs</text>';
    return s + "</svg>";
  }

  // ---- wire up -----------------------------------------------------------
  MFD.forEach(function(d){ buildDial(d, $("pca-mfd"), "mfd"); });
  NAT.forEach(function(d){ buildDial(d, $("pca-nat"), "nat"); });
  $("pca-resim").addEventListener("click", function(){
    $("pca-seed").value = (parseInt($("pca-seed").value, 10) || 1) + 1;
    resimulate();
  });
  $("pca-seed").addEventListener("change", resimulate);
  $("pca-reset").addEventListener("click", function(){
    S = Object.assign({}, DEFAULTS); lastF = DEFAULTS.f; lastB = DEFAULTS.b;
    resimulate();
  });
  resimulate();
})();
</script>
```

## Things to try

**Turn `f` up towards 1.** Both excess rates fall to zero and the two bars
converge — the sources agree completely. Now notice what happens to $\beta$: it
stops meaning anything. There is no excess left to be biased, so the bias is
*unidentified*, and the widget flags it and holds the last value rather than
snapping to zero. Zero would be a claim (the methods are unbiased) that the
state cannot support.

**Drag `lambdaem` alone.** $M$, $f$ and $\beta$ all move, because changing one
excess rate changes the overall level, the shared share, *and* the imbalance
simultaneously. This is exactly why the native coordinates are awkward to
reason in, and it is much easier to see than to describe.

**Turn $\beta$ up and watch $M$.** The two bars separate, but the $M$ rule
does not move — it stays at their *average*, touching neither. The average of
the two excess rates is $M(1-f)$ for any $\beta$, so $M$ is pinned to the
midpoint whatever the bias. $M$ is a midpoint of what the sources *report*, not
a property of the underlying process.

**Compare $\kappa_A$ against $\kappa_X$.** They are near-orthogonal channels.
$\kappa_A$ moves the pair up and down *together* — it cannot touch the
difference, because the shared component cancels from it. $\kappa_X$ pulls the
pair *apart*, and drives the whole difference.

**Push $\kappa_A$ to 0.** The counts do not stop moving. That is the Poisson
floor, not determinism: there is no parameter setting that stills the stacks.

## The same map, in R

The widget's arithmetic is not a separate model — it is the package's own map,
which you can call directly. This is the authoritative version, and it is what
the JavaScript above is checked against:

```{r}
#| label: map
library(pairedcountbrms)

# The widget's defaults
binegbin_mfd_to_dpars(M = 12, f = 0.67, delta = atanh(0), kappas = 0.6, kappax = 1.0)
```

`delta` is the unbounded log-ratio bias; the widget's $\beta$ is the bounded
$\tanh\delta$. Going back the other way:

```{r}
#| label: inverse
d <- binegbin_mfd_to_dpars(M = 12, f = 0.67, delta = 0.3)
binegbin_dpars_to_mfd(d$mu, d$lambdaem, d$lambdalb)[c("M", "f", "delta", "beta")]
```

And the degenerate case the widget flags:

```{r}
#| label: degenerate
# Perfect congruence: no excess, so no bias to identify
binegbin_dpars_to_mfd(mu = 12, lambdaem = 0, lambdalb = 0)$delta
```

## Fitting in these coordinates

Nothing above fits anything — these are coordinate transforms. To *fit* in
$(M, f, \delta)$ you do not need a different family, because the
reparameterisation is reachable through a non-linear formula. All five dpars are
log-linked, so the link supplies the `exp()`:

```{r}
#| label: nlf
#| eval: false
bf(y_em | vint(y_lb) ~ 1, nl = TRUE) +
  nlf(mu       ~ eta + log_inv_logit(con)) +
  nlf(lambdaem ~ log(2) + eta + log_inv_logit(-con) + log_inv_logit( 2 * methd)) +
  nlf(lambdalb ~ log(2) + eta + log_inv_logit(-con) + log_inv_logit(-2 * methd)) +
  lf(eta ~ 1, con ~ 1, methd ~ 1)
```

with `eta` $= \log M$, `con` $= \operatorname{logit} f$, and `methd` $= \delta$.
Putting a random effect on `eta` asks whether groups differ in overall level;
putting one on `con` asks whether they differ in congruence. Those are separable
questions in these coordinates and entangled ones in the native dpars, which is
the practical reason to bother.

## A caveat about the widget

The JavaScript is an independent implementation of the *generative* model — it
draws from the same trivariate reduction, but it is not the likelihood, and no
part of the package's inference runs in your browser. The rate map it uses is
the one tested in `test-mfd.R`; the sampler is illustrative. Treat the picture
as intuition, and the R above as the specification.
